The 2018 heatwave and its implications on ozone induced damage on vegetation in a subarctic climate

### 2020-03-27 VegOzone.tex
Fixed broken reference.
Ane - added text about stomata and ozone.

\introduction  %% \introduction[modified heading if necessary]
\label{sec:intro}
\begin{itemize}
\item ozone in general
\item ozone induced damage in vegetation \emph{Contributor: Ane - maybe also ask Aud?}
\item present subarctic climate and conditions
  \begin{itemize}
  \item[$\rightarrow$] midnight sun and 23\,\unit{h} open stomata (biological details will be published in a component paper)
  \item[$\rightarrow$] ozone climatology and spring peak
  \end{itemize}
\item future climate and conditions
  \begin{itemize}
  \item[$\rightarrow$] increase in heatwaves and dry periods
  \item[$\rightarrow$] shift in start of growing season (into vicinity of ozone spring peak)
  \item[$\rightarrow$] increase in tropospheric background ozone
  \end{itemize}
\item meteorological situation 2018 vs 2019
\end{itemize}

@@

In our assessment on ozone damage on vegetation in the subarctic in relation to the 2018 heatwave, we focus on the NIBIO operated local site Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley (Section~\ref{subsec:svanhovd}). Throughout the paper, we will use Svanvik synonymously for the atmospheric monitoring site at Svanhovd. The available agrameteorological monitoring data are described in Section~\ref{subsec:atmo_svanvik}. In Section~\ref{subsubsec:ozone_garden}, the ozone garden -- a plant-based ozone monitoring -- will be described in more detail. For evaluation of the ozone data taken at Svanhovd, we compare with observational data from other monitoring sites in northern Fennoscandia (Section~\ref{subsec:obs_data}). Reanalysis data is used as complement where observational data was not available or temporally sparse. The chosen data sets are described in Section~\ref{subsec:model_data}.

>>>>

\introduction  %% \introduction[modified heading if necessary]
\label{sec:intro}
\begin{itemize}
\item ozone in general
\item ozone induced damage in vegetation \emph{Contributor: Ane - maybe also ask Aud?}
  Ozone acts as an oxidative stress to plants. Its main action is imposed through reactions occuring in the cell walls and cell membranes of mesophyll cells inside the leaves.Ozone enters the leaves through stomata, that are kept open to allow for gas exchange between the leaf air spaces and the exterior air. When the stomata are closed, ozone does not enter.
\item present subarctic climate and conditions
  \begin{itemize}
  \item[$\rightarrow$] midnight sun and 23\,\unit{h} open stomata (biological details will be published in a component paper)
  \item[$\rightarrow$] ozone climatology and spring peak
  \end{itemize}
\item future climate and conditions
  \begin{itemize}
  \item[$\rightarrow$] increase in heatwaves and dry periods
  \item[$\rightarrow$] shift in start of growing season (into vicinity of ozone spring peak)
  \item[$\rightarrow$] increase in tropospheric background ozone
  \end{itemize}
\item meteorological situation 2018 vs 2019
\end{itemize}

@@

In our assessment on ozone damage on vegetation in the subarctic in relation to the 2018 heatwave, we focus on the NIBIO operated local site Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley (Section~\ref{subsec:svanhovd}). Throughout the paper, we will use Svanvik synonymously for the atmospheric monitoring site at Svanhovd. The available agrameteorological monitoring data are described in Section~\ref{subsubsec:atmo_svanvik}. In Section~\ref{subsubsec:ozone_garden}, the ozone garden -- a plant-based ozone monitoring -- will be described in more detail. For evaluation of the ozone data taken at Svanhovd, we compare with observational data from other monitoring sites in northern Fennoscandia (Section~\ref{subsec:obs_data}). Reanalysis data is used as complement where observational data was not available or temporally sparse. The chosen data sets are described in Section~\ref{subsec:model_data}.
